### 2021-02-22 text/draft.tex
reconcilement

Machine indexing of institutional repositories: indexing Edoc with Annif
and BARTOC FAST as proof of concept

@@

In second step, a list of all single keywords must be created. In order
to do so, let us consider now in more detail the exact information
extracted from the \texttt{keywords} data fields as per
\texttt{keywords/keywords\_extracted.json}. In Edoc, the
\texttt{keywords} data field of an item is a non-mandatory free text
field that is filled in by the user (usually one of the authors) who
undertakes the data entry of an item to Edoc. Even though the Edoc user
manual specifies that keywords must be separated by commas (Universität
Basel, n.d., 8), this requirement is neither validated by the input mask
nor by an administrator of Edoc. Furthermore, neither the manual nor the
input mask provide a definition of the term ``keyword.'' A vocabulary or
a list of vocabularies from which to choose the keywords is also
lacking. Taken together, these observations are indicative of very
heterogeneous data in the \texttt{keywords} data field. To wit, the
items of \texttt{keywords/keywords\_extracted.json} are strings where
single keywords are individuated by any symbols the user saw fit. So,
for each item in \texttt{keywords/keywords\_extracted.json}, the user
input must be parsed into single keywords.

@@

\texttt{\_Keywords.clean\_keywords} is the implementation the parser
while the recursive cleaner is implemented by
\texttt{\_Keywords.clean\_keyword}. To create the desired list of clean
keywords, saved as \texttt{keywords/keywords\_clean.json}, we call
\texttt{\_Keywords.clean\_keywords} with the extracted keywords as
argument:

@@

\texttt{keywords/keywords\_clean.json} has 36'901 entries many of which
are duplicates. We hence deduplicate and count the number of occurrences
of each keyword. This is achieved by calling
\texttt{\_Keywords.make\_histogram} on
\texttt{keywords/keywords\_clean.json} and saving the output as
\texttt{keywords/keywords\_clean\_histogram.json}:

@@

An analysis of \texttt{keywords/keywords\_clean\_histogram.json} shows
that the lion's share of keywords has only one occurrence but that the
total occurrences are predominantly made up of keywords with more than
one occurrence (see Figure 3 for details).

@@

The data from \texttt{keywords/keywords\_clean\_histogram.json} was
imported into OpenRefine. The \texttt{keyword} column was then
duplicated and reconciled with Wikidata. Here the parameters were chosen
as follows: reconcile against no particular type, and auto-match
candidates with high confidence.

There were X automatic matches (call them ``suggestions''). A suggestion
is correct if and only if the meaning of the keyword from
\texttt{keywords/keywords\_clean\_histogram.json} corresponds to the
meaning of the suggested concept from Wikidata. Note that for homonymous
or polysemous keywords, it is impossible to confirm a correct match
without further context; those keywords therefore cannot be reconciled
(but see \protect\hyperlink{construction}{section ``Construction''} for
a possible solution). Unfortunately, random sampling showed that the
overall quality of the reconciliation was not satisfactory, that is,
there were too many incorrect suggestions. A two-pronged strategy was
adopted to ameliorate the quality of the reconciliation.

@@

With these improvements in place, each keyword with a suggestion was
assigned a QID via the ``Add cloumns from reconciled values''-function
(and similar for YSO and MeSH identifiers). The data was then exported
and saved as \texttt{/keywords/keywords\_reference.json}. Keywords with
a QID now constitute 69\% of all keywords and 78\% of their total
occurrences in the sample data set, but these numbers are much lower for
YSO and MeSH identifiers (see Figure 4 for details).

\begin{figure}
\centering
\includegraphics{images/native_gold_standard.pdf}
\caption{The coverage of \texttt{/keywords/keywords\_reference.json} by
the controlled vocabularies of Wikidata (QID), Medical Subject Headings
(MeSH), and YSO (General Finnish Ontology). Both MeSH and YSO are
dependent on QID but independent of each other.}
\end{figure}

Let us now turn to the evaluation of the reconciliation: how many of the
suggestions were correct? This question was answered via random
sampling. The random sample (\(n=500\)) was created by calling
\texttt{\_Analysis.make\_random\_sample} on
\texttt{/keywords/reference\_keywords.json}; it is available at
\texttt{analysis/random\_keywords.json}. The sample was then imported
into OpenRefine and the judgement (1 for correct, 0 for incorrect) of
the manual verification was recorded in the column
\texttt{verification}. The data was then exported and is available at
\texttt{analysis/random\_keywords\_verified.csv}.

@@

!! Talk about enrichment with yso and mesh

@@

!! Todo: - Explain MeSH\\
- Explain how to get MeSH for subset of sample data set

>>>>

@@

In second step, a list of all single keywords must be created. In order
to do so, let us consider now in more detail the exact information
extracted from the \texttt{keywords} data fields as per
\texttt{/keywords/keywords\_extracted.json}. In Edoc, the
\texttt{keywords} data field of an item is a non-mandatory free text
field that is filled in by the user (usually one of the authors) who
undertakes the data entry of an item to Edoc. Even though the Edoc user
manual specifies that keywords must be separated by commas (Universität
Basel, n.d., 8), this requirement is neither validated by the input mask
nor by an administrator of Edoc. Furthermore, neither the manual nor the
input mask provide a definition of the term ``keyword.'' A vocabulary or
a list of vocabularies from which to choose the keywords is also
lacking. Taken together, these observations are indicative of very
heterogeneous data in the \texttt{keywords} data field. To wit, the
items of \texttt{/keywords/keywords\_extracted.json} are strings where
single keywords are individuated by any symbols the user saw fit. So,
for each item in \texttt{/keywords/keywords\_extracted.json}, the user
input must be parsed into single keywords.

@@

\texttt{\_Keywords.clean\_keywords} is the implementation the parser
while the recursive cleaner is implemented by
\texttt{\_Keywords.clean\_keyword}. To create the desired list of clean
keywords, saved as \texttt{/keywords/keywords\_clean.json}, we call
\texttt{\_Keywords.clean\_keywords} with the extracted keywords as
argument:

@@

\texttt{/keywords/keywords\_clean.json} has 36'901 entries many of which
are duplicates. We hence deduplicate and count the number of occurrences
of each keyword. This is achieved by calling
\texttt{\_Keywords.make\_histogram} on
\texttt{/keywords/keywords\_clean.json} and saving the output as
\texttt{/keywords/keywords\_clean\_histogram.json}:

@@

An analysis of \texttt{/keywords/keywords\_clean\_histogram.json} shows
that the lion's share of keywords has only one occurrence but that the
total occurrences are predominantly made up of keywords with more than
one occurrence (see Figure 3 for details).

@@

The data from \texttt{/keywords/keywords\_clean\_histogram.json} was
imported into OpenRefine. The \texttt{keyword} column was then
duplicated and reconciled with Wikidata. Here the parameters were chosen
as follows: reconcile against no particular type, and auto-match
candidates with high confidence.

There were X automatic matches (call them ``suggestions''). A suggestion
is correct if and only if the meaning of the keyword from
\texttt{/keywords/keywords\_clean\_histogram.json} corresponds to the
meaning of the suggested concept from Wikidata. Note that for homonymous
or polysemous keywords, it is impossible to confirm a correct match
without further context; those keywords therefore cannot be reconciled
(but see \protect\hyperlink{construction}{section ``Construction''} for
a possible solution). Unfortunately, random sampling showed that the
overall quality of the reconciliation was not satisfactory, that is,
there were too many incorrect suggestions. A two-pronged strategy was
adopted to ameliorate the quality of the reconciliation.

@@

With these improvements in place, each keyword with a suggestion was
assigned a QID via the ``Add cloumns from reconciled values''-function
(and similar for YSO and MeSH identifiers). The data was then exported
and saved as \texttt{/keywords/keywords\_reference.json}. Keywords with
a QID now constitute 69\% of all keywords and 78\% of their total
occurrences in the sample data set, but these numbers are significantly
lower for the MeSH identifier (53.6\% of all keywords with only 28.8\%
of all occurrences) and especially low for the YSO identifier (X of all
keywords with Y of all occurrences). In both cases, the problem is due
to the fact that Wikidata's mapping of MeSH respectively YSO is only
partial. There can be two reasons for this state of affairs for a given
entry: either there is no match between Wikidata and YSO or MeSH (after
all, Wikidata is much larger than MeSH and YSO taken together), or there
is a match but it has not yet been added to the mapping. In the latter
case, at least with respect to YSO, there is a solution: Finto provides
a REST-sytle API to access the YSO vocabulary directly (see
https://api.finto.fi/). For each keyword in in the list of reference
keywords that lacks a YSO identifier, the Finto API is queried; if a
term turns up, it is added to the keyword. The effect of this second
reconcilement is detailed in Figure 4. It is achieved by calling
\texttt{\_Keywords.enrich\_with\_yso} on
\texttt{/keywords/keywords\_reference.json} and saving the output as
\texttt{/keywords/keywords\_reference\_master.json}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_reference }\OperatorTok{=}\NormalTok{ \_Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/keywords/keywords\_reference.json"}\NormalTok{)}
\NormalTok{keywords\_reference\_new }\OperatorTok{=}\NormalTok{ \_Keywords.enrich\_with\_yso(keywords\_reference)}
\NormalTok{\_Utility.save\_json(keywords\_histogram, DIR }\OperatorTok{+} \StringTok{"keywords/keywords\_reference\_master.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{images/native_gold_standard.pdf}
\caption{The coverage of
\texttt{/keywords/keywords\_reference\_master.json} by the controlled
vocabularies of Wikidata (QID), Medical Subject Headings (MeSH), and YSO
(General Finnish Ontology). Both MeSH and YSO are dependent on QID but
independent of each other. YSO enriched is the superset of YSO created
by reconciling keywords that lack a YSO identifier directly with the YSO
database provided by the Finto API; it is hence independent of
Wikidata.}
\end{figure}

Let us now turn to the evaluation of the reconciliation: how many of the
suggestions were correct? This question was answered via random
sampling. The random sample (\(n=500\)) was created by calling
\texttt{\_Analysis.make\_random\_sample} on
\texttt{/keywords/reference\_keywords\_master.json}; it is available at
\texttt{/analysis/random\_keywords.json}. The sample was then imported
into OpenRefine and the judgement (1 for correct, 0 for incorrect) of
the manual verification was recorded in the column
\texttt{verification}. The data was then exported and is available at
\texttt{/analysis/random\_keywords\_verified.csv}.

@@

@@

\begin{itemize}
\tightlist
\item
  !! Explain MeSH
\item
  !! Explain how to get MeSH for subset of sample data set
\end{itemize}
